---
title: 'Laboratory Exercise Week 3'
author: "Your Name and Section, 10 pts"
date: "Todays Date"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed laboratory exercise using Blackboard's Turnitin feature. Your Turnitin upload link is found on your Blackboard Course shell under the Laboratory folder.

***

For this exercise, you will need to use the package `mosaic` to find numerical and graphical summaries.


```{r warning=FALSE, message=FALSE}
# install mosaic package if necessary
if (!require(mosaic)) install.packages(`mosaic`)
# load the package in R
library(mosaic) # load the package mosaic to use its functions
```


1. Recall the `iris` data set from last week's exercise. The `iris` data set is already pre-loaded in R - look at the help file using `?iris` for more information on this data set.   
      i) Check the structure of the data using the function `str(iris)`.   
      ii) Find the average (or mean) measurement of the variable `Sepal.Length`. Do this in two ways as described in the lesson.   
      iii) Find the average `Sepal.Length` for the different flower `Species`. Give a brief comment on the averages.  
      iv) Repeat (ii) and (iii) but use the summary standard deviation `sd()` which describes the spread of the variable.   
      v) Describe the shape of the variable `Sepal.Length` by creating a histogram using `histogram()`.  Write your description outside the code chunk.  
      vi) Compare the `Sepal.Length` of the three species of flowers by creating a side-by-side boxplot using `bwplot()`. Write your description outside the code chunk.
      

### Code chunk
```{r} 
# Insert your code for this question after this line	
str(iris)
mean(~Sepal.Length, data = iris)
mean(iris$Sepal.Length)
mean(Sepal.Length~Species, data = iris) #Setosa has the lowest average sepal length, followed by versicolor, then virginica.
sd(~Sepal.Length, data = iris)
sd(iris$Sepal.Length)
histogram(iris$Sepal.Length) #looks fairly normaly distributed with a slight skewness to the left
bwplot(Species~Sepal.Length, data = iris) # fairly differing sepal lengths within the different species


# last R code line
```


2. The data set `MLB-TeamBatting-S16.csv` contains MLB Team Batting Data for selected variables. Load the data set from the given url using the code below. This data set was obtained from [Baseball Reference](https://www.baseball-reference.com/leagues/MLB/2016-standard-batting.shtml).
    * Tm - Team   
    * Lg - League: American League (AL), National League (NL)  
    * BatAge - Batters’ average age  
    * RPG - Runs Scored Per Game  
    * G - Games Played or Pitched  
    * AB - At Bats  
    * R - Runs Scored/Allowed  
    * H - Hits/Hits Allowed  
    * HR - Home Runs Hit/Allowed  
    * RBI - Runs Batted In  
    * SO - Strikeouts  
    * BA - Hits/At Bats  
    * SH - Sacrifice Hits (Sacrifice Bunts)  
    * SF - Sacrifice Flies  

    i) Find the average measurement for the following variables `BatAge`, `RPG`, `R`, `H` and `BA`.   
    ii) Create dotplot's or histogram's for each variable in (i).     
    iii) Using your own words, describe the distribution of each variable in (i).   Write your answer outside the code chunk.     
    iv) Find the average and the standard deviation of the variables `RPG`, `H` and `BA` for each league.     
    v) Describe any differences or similarities between the leagues. Write your comment outside the code chunk.
      


### Code chunk
```{r} 
# load the data set
mlb16.data <- read.csv("https://raw.githubusercontent.com/jpailden/rstatlab/master/data/MLB-TeamBatting-S16.csv")
str(mlb16.data) # check structure
head(mlb16.data)  # show first six rows

mean(mlb16.data$BatAge)
mean(mlb16.data$RPG)
mean(mlb16.data$R)
mean(mlb16.data$H)
mean(mlb16.data$BA)	

histogram(~BatAge, data = mlb16.data) #normally distributed with a slight skewness to the left
histogram(~RPG, data = mlb16.data) #almost perfect normal distribution	
histogram(~R, data = mlb16.data) #almost perfect normal destribution within the 720 range with a slight skewness to the left
histogram(~H, data = mlb16.data) #large skewness to the left
histogram(~BA, data = mlb16.data) #a uniform normal distribution with a slight skewness to the left and a low mean

favstats(RPG~Lg, data = mlb16.data) #AL has a higher mean compared to NL and the min and max with a lower standard deviation. This would indicate that the players are overall scoring more runs per game and with higher consistency
favstats(H~Lg, data = mlb16.data) #the teams in AL have allowed more hits compared to NL, which is evident by a higher mean, min, and max. The standard deviation is also lower compared to NL. 
favstats(BA~Lg, data = mlb16.data) #Both leagues have relatively similar battering averages in terms of mean and standard deviation
#with a higher number of hits allowed and with the same number of hits at bat, it makes sense that AL has a higher amount of runs per game compared to NL.
# last R code line

```



